### 2012-10-24 rnaseq_paper.tex
edited the introduction

\section*{Introduction}
Until recently, studies of alternative splicing had been limited to a small number of genes and isoforms due to high-cost and low-throughput sequencing of expressed sequence tags (ESTs) and full-length cDNA libraries.  
RNA sequencing (RNA-Seq) using the next-generation sequencing (NGS) has been used successfully in many studies to gain unprecedented insight into a complexity of transcriptomes.
Novel isoforms have been identified based on reads mapped across known splice junctions or putative splice junctions from annotations.
It has been estimated that, in human, $92-94$\% of multiexon genes undergo alternative splicing and different isoforms are expressed in different tissues\cite{Wang:2008ea}.
This suggests that even in human a large number of splice variants have not been explored.

Sequences from RNA-Seq are very short (~75--250bp for Illumina reads); therefore, it is not feasible to assess gene
and isoform expression without mapping reads to a reference annotation or construct a full-length transcript from
\emph{de novo} assembly.
Several softwares employing different approaches have been developed to reconstruct transcripts from short reads.
Tophat\cite{Trapnell:2009dp} maps reads to a reference genome to identify exons and splice sites \emph{ab initio}.
It first searches for known splie sites in a genome and then tries to map reads across splice juntions to identify exon junctions.
Exon junctions can be used to build gene models by Cufflinks\cite{Trapnell:2010kd}.
On the other hand, Trinity\cite{Grabherr:2011jb}, has been developed to assemble short reads \emph{de novo} to construct transcripts.
Trans-Abyss\cite{Robertson:2010ih} and Oases\cite{Schulz:2012je} are an extension of Abyss\cite{Simpson:2009iv} and Velvet\cite{Zerbino:2008vu,Zerbino:2009jp} assembler that are designed to work with RNA-Seq data.
These programs are comparable at identifying alternative splicing with a slightly different sensitivity and specificity.
However, Oases with Oases-M has been shown to be superior than other \emph{de novo} assembler at discovering isoforms in human and mouse\cite{Schulz:2012je}.

In this study, we present a pipeline that uses Velvet and Oases assembler to discover alternative splicing in RNA-Seq data from chickens.
We use a new technique called a local assembly that enhances sensitivity of alternative splicing detection of Oases.
The results show that the pipeline can detect more isoforms than Oases-M and can detect isoforms not found by
Cufflinks.

>>>>

\section*{Introduction}
Until recently, studies of alternative splicing had been limited to a small
number of genes and isoforms due to high-cost and low-throughput sequencing
of expressed sequence tags (ESTs) and full-length cDNA libraries.  
RNA sequencing (RNA-Seq) using the next-generation sequencing (NGS) has been
used successfully in many studies to gain unprecedented insight into a
complexity of transcriptomes.
It has been estimated that, in human, $92-94$\% of multiexon genes undergo alternative
splicing and different isoforms are expressed in different tissues\cite{Wang:2008ea}.
This suggests that even in human a large number of splice variants have not been explored.

Since sequences from RNA-Seq (reads) are very short (~50--250bp for Illumina reads),
it is not feasible to assess expression isoforms and their structures without mapping reads
to a reference annotation or reconstructing a full-length transcript from short reads.

Despite the small size of sequencing reads, several studies have detected
novel splice junctions based on alignment of reads spanning across putative exons.
To map reads across splice junctions, reads are split into two parts and mapped
to the genome independently.
A splice junction is then identified based on alignments of each half of a read
that fall between two exons at exon-intron boundaries.
With this approach, Wang \emph{et al.} have identified a large number of
splice junctions that are not annotated from human cell lines.
These novel splice sites include both canonical splie sites and non-canonical splice site.
Approximately, $46\%-75\%$ of canonical splice sites are supported by ESTs.
Splice junctions have different read coverage, which suggests that both highly
and lowly expressed isoforms are not annotated.

Pickrell \emph{et al.} have also identified more than 150,000 novel canonical splice junctions
in lymphoblastoid cells using a similar approach.
The study has also shown that the number of unannotated splice junctions varies among cells from
different human tissues, which suggests tissue-specific expression of isoforms.
A majority of unannotated isoforms found in both studies are from alternative splicing
of annotated exons a small fraction of isoforms contain exons from intergenic or intronic region.

Several tools have been developed not only to detect novel splice junctions but also to reconstruct
full-length isoforms from short reads with or without gene annotations.
Cufflinks\cite{Trapnell:2010kd} relies on splice junctions detected from Tophat\cite{Trapnell:2009dp},
a read aligner that can align reads across putative exons, to reconstruct a full-length transcripts.
Cufflinks was used to identified 12,712 novel isoforms, of which 7,395 ($58\%$) contain novel splice
junctions in mouse myoblast cell lines.
The mapping-based approach has been useful in both detecting splice junctions and isoforms.
One downside of this approach is it relies heavily on a reference genome.
Therefore, organisms lacking a genome or a high-quality one may not find this approach practical.
This limitation can be circumvent by employing \emph{de novo} assembly approach.

A number of \emph{de novo} assemblers have been used to reconstruct transcripts from RNA-Seq data in
many studies.
Trinity\cite{Grabherr:2011jb}, has successfully been used to reconstruct transcripts both yeast and mouse.
It has also shown that Trinity detects unique sets of splice junctions when compared to annotated
junctions and junctions detected by Cufflinks.
This suggests that \emph{de novo} assembly approach is capabale of increasing sensitivity of detecting
alternative isoforms.
Trans-Abyss\cite{Robertson:2010ih} and Oases\cite{Schulz:2012je} are an extension
of Abyss\cite{Simpson:2009iv} and Velvet\cite{Zerbino:2008vu,Zerbino:2009jp} assembler
that are tuned to work with RNA-Seq data.
These programs are comparable at reconstructing known and novel alternative isoforms with a slightly
different sensitivity and specificity.
However, Oases with Oases-M has been shown to be superior than other \emph{de novo}
assembler at discovering isoforms in human and mouse\cite{Schulz:2012je}.

In this study, we present a pipeline that uses Velvet and Oases assembler to
reconstruct alternative isoforms in RNA-Seq data from chickens.
We use a new technique called a local assembly that enhances sensitivity of
alternative splicing detection of Oases.
The results show that the pipeline can detect more isoforms than Oases-M and
can detect isoforms not found by Cufflinks.
